### 2015-06-18 inst/examples/plot_prelim_mcmc.Rmd
adds plot of prelim results example

---
title: ''
output: html_document
---

Check preliminary results from Rstan MCMC. Took 19 days and 9 hours to fit the full spatial dataset at 2000 iterations with 3 chains in parallel.

```{r plot_mcmc, echo=TRUE, include=TRUE, message=FALSE, warning=FALSE}
library(ggmcmc)
library(ggplot2)
mcmc <- readRDS("../../results/stanmcmc_sage.RDS")
ggplot(subset(mcmc, Parameter=="beta_mu"),
       aes(x=Iteration, y=value, color=as.factor(Chain)))+
  geom_line()
ggplot(subset(mcmc, Parameter=="int_mu"),
       aes(x=Iteration, y=value, color=as.factor(Chain)))+
  geom_line()
ggplot(subset(mcmc, Parameter=="alpha[1]"),
       aes(x=Iteration, y=value, color=as.factor(Chain)))+
  geom_line()
```

The slope mixed well, but clearly the alphas and the intercept are fighting. Need to work this out with Mevin.