### 2015-11-03 docs/appendix_a.Rmd
updates model table

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
bibliography: ~/Dropbox/Bibliography/SageAbundance.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: true
fontsize: 12pt
geometry: margin=1in

---

\renewcommand\thefigure{X\arabic{figure}}

Appendix X: Table of CMIP5 models used in this study
=======================================

\scriptsize

| Modeling Center | Model | Institution |
| --------------- | ----- | ----------- |
| BCC | BCC-CSM1.1 | Beijing Climate Center, China Meteorological Administration | 
| BCC | BCC-CSM1.1(m) | Beijing Climate Center, China Meteorological Administration | 
| CCCma | CanAM4 | Canadian Centre for Climate Modelling and Analysis |
| CCCma | CanCM4 | Canadian Centre for Climate Modelling and Analysis |
| CCCma | CanESM2 | Canadian Centre for Climate Modelling and Analysis |
| CCCMC | CMCC-CESM | Centro Euro-Mediterraneo per I Cambiamenti Climatici |
| CCCMC | CMCC-CM | Centro Euro-Mediterraneo per I Cambiamenti Climatici |
| CCCMC | CMCC-CMS | Centro Euro-Mediterraneo per I Cambiamenti Climatici |
| CNRM-CERFACS | CNRM-CM5 | Centre National de Recherches Meteorologiques / |
|  |  | Centre Europeen de Recherche et Formation Avancees en Calcul Scientifique |
| CNRM-CERFACS | CNRM-CM5-2 | Centre National de Recherches Meteorologiques / |
|  |  | Centre Europeen de Recherche et Formation Avancees en Calcul Scientifique |
| COLA and NCEP | CFSv2-2011 | Center for Ocean-Land-Atmosphere Studies and |
|  |  | National Centers for Environmental Prediction |
| CSIRO-BOM | ACCESS1.0 | CSIRO (Commonwealth Scientific and Industrial Research |
|  |  | Organisation, Australia), and BOM (Bureau of Meteorology, Australia) |
| CSIRO-BOM | ACCESS1.3 | CSIRO (Commonwealth Scientific and Industrial Research |
|  |  | Organisation, Australia), and BOM (Bureau of Meteorology, Australia) |