### 2015-11-03 docs/appendixB.Rmd
updates knot appendix B

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
bibliography: ~/Dropbox/Bibliography/SageAbundance.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
fontsize: 12pt
geometry: margin=1in

---

\renewcommand\thefigure{B\arabic{figure}}

Appendix B: Detailed methods for knot choice and placement
==========================================================
Our dimension reduction approach requires fitting random effects associated with knots distributed nonrandomly in space. By distributing *m* knots across space, we can fit random effects for each knot and then use process convolution to interpolate the random effect between the knots. Increasing the number of knots results in a more finely-resolved spatial effect, but it also increases processe time. Thus, there exists a tension between the resolution of the spatial effect and computational effort. Without regard to computation time, it is possible to use common model selection techniques to guide the number of knots one should use. With regard to computation time, it is sensible to choose knot placement and number based on the observed residual spatial variation.

To do so, we fit a generalized linear model of the form

\vspace{-4em}
\begin{align}
y_{i,t} \sim \text{Poisson}(\mu_{i,t}),
\end{align}
\vspace{-4em}

where $\mu_{i,t}$ is the expected percent cover of pixel *i* in year *t*

\vspace{-4em}
\begin{align}
\text{log}(\mu_{i,t}) &= \beta_{0} + \beta_{1}y_{i,t-1} + \textbf{x}_{t}'\boldsymbol{\gamma}
\end{align}
\vspace{-3em}

that includes a density-dependence effect of log-transformed cover in the previous year ($y_{i,t-1}$) and climate effects ($\textbf{x}_t$). In other words, a non-hierarchial version of the model presented in our main text. We then examined a spatial variogram of the residuals from the above model to identify the the distance at which spatial dependence disappears.

Spatial dependence existed in the residuals until pixels were separated by about 500 meters (Fig. B1). Thus, we set the kernal bandwidth ($\sigma$ in Eq. X of the main text) to 500/3 = `r round(500/3,2)`. We then placed equally-spaced knots across our spatial grid until the distance between nearest neighbors was approximately $\sigma$ = `r round(500/3,2)`. This resulted in 121 knots (Fig. B2).

![Spatial variogram of model residuals showing where spatial dependence disappears (dark red line at ~500 meters), and the value chosen for the kernal bandwidth in Eq. X of the main text (500/3; dashed black line).](components/avgresidual_variogram_withrange_subset.png)

![Map of our study area showing percent cover in 1984 and the location of the 121 evenly-spaced knots.](components/SAGE_Grid_wKnots_subset.png)