### 2015-11-30 docs/appendixD.Rmd
updates appendices

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
bibliography: ~/Dropbox/Bibliography/SageAbundance.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
fontsize: 12pt
geometry: margin=1in

---

\renewcommand\thefigure{D\arabic{figure}}

Appendix D: Posterior estimate of spatial random effects ($\boldsymbol{\eta}$)
==============================================================================

```{r figure_D1,dependson="plot-options", fig.cap="Standardized percent cover showing the general spatial structure of above and below average pixels in the data (top panel) compared to the mean posterior estimate of the spatial field, $\\eta$ (bottom panel).", eval=TRUE, echo=FALSE}
library(png)
library(grid)
img <- readPNG("../results/spatialfield.png")
grid.raster(img)
```