### 2015-12-30 docs/appendixE.Rmd
adds app E with forecast densities

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
bibliography: ~/Dropbox/Bibliography/SageAbundance.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
fontsize: 12pt
geometry: margin=1in

---

\renewcommand\thefigure{E\arabic{figure}}

Appendix E: Density plots of sagebrush cover forecasts ($\boldsymbol{\eta}$)
==============================================================================

```{r figure_E1,dependson="plot-options", fig.cap="Estimated kernel densities for current and projected sagebrush cover across space and simulation time. These densities represent the desnity of cover values from the spatial plots in Figure 5 of the main text.", eval=TRUE, echo=FALSE}
library(png)
library(grid)
img <- readPNG("../results/clim_change_densities.png")
grid.raster(img)
```