### 2016-06-17 docs/appendixB.Rmd
adds new appendix; includes ms revisions; ready for coauthors

---
layout: 12pt
header-includes:
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{graphicx}
bibliography: ~/Dropbox/Bibliography/SageAbundance.bib
csl: components/ecology.csl

## rmarkdown render options
output:
  pdf_document:
    fig_caption: true
    keep_tex: false
fontsize: 12pt
geometry: margin=1in

---

\renewcommand\thefigure{B\arabic{figure}}

Appendix B: Annual maps of sagebrush cover
==========================================

We used remotely-sensed estimates of sagebrush cover [@Homer2012].
Here we show the annual maps of sagebrush cover as hindcasted with the regression tree algorithm described in @Homer2012.

\begin{figure}[!ht]
  \centering
      \includegraphics[width=7in]{../figures/all_years_percCover.png}
  \caption{Estimated percent cover of 30 x 30 meter pixels in our study area in southwestern Wyoming (see Fig. 1). Maps show percent cover for each year (1984-2011).}
\end{figure}